---
title: Simulation
                 \usepackage{float}
header-includes: \usepackage{color}
output:
  html_document: default
  pdf_document:
    fig_caption: no
---
<style>
table, th, td { text-align:right; }
th, td {padding: 10px;}
</style>

```{r, echo=FALSE, warning=FALSE, message=FALSE}
whichcomp <- strsplit(getwd(),"/")[[1]][3]
load(paste0("c:/users/", whichcomp, "/Dropbox/teaching/Resma3/Resma3.RData"))
library(knitr)
library(wolfr)
opts_chunk$set(fig.width=6, fig.align = "center", 
      out.width = "70%", warning=FALSE, message=FALSE)
library(ggplot2)
library(grid)
```
`r hl()$basefontsize()`

A simulation is a way to do experiments on a computer. This can be useful to do various calculations as well as to test the performance of different methods.  

A simulation generally consists of the following parts:

- generate random data 
- calculate something for this data  
- repeat the above many times, keep track of results
- analyze those results

**Example:** say we are interested in the following: if a fair die is rolled, what is the probability of a six?  

Now we already know the answer (1/6) but let's do a simulation anyway.

- generate random data    
in the real live experiment the data are the numbers 1-6, each with probability 1/6. We can generate data like this as follows:

```{r}
x <- sample(1:6, size = 20, replace = TRUE)
x
```

- calculate something for this data  

we want to know if x is 6 or not. This we can do with

```{r}
x==6
```

- repeat the above many times, keep track of results  

in the calculations above we already did the simulation 20 times. 

In general one wants to do this 10000 times (or so). Of course then we can't print all the stuff to the screen.

```{r}
B <- 10000
x <- sample(1:6, size = B, replace = TRUE)
```

- analyze those results  

we want to know the proportion of 6's, so

```{r}
sum(x==6)/B
```

Note that each time we run a simulation the computer generates new data, so the answer will always be a little bit different:

```{r}
sum(sample(1:6, size = B, replace = TRUE)==6)/B
sum(sample(1:6, size = B, replace = TRUE)==6)/B
sum(sample(1:6, size = B, replace = TRUE)==6)/B
```

`r hl()$vspace()`


## Generating Data

R has a great many standard probability distributions built in. The general format is **r** followed by the name (sort of). So for example to generate 20 observations  from a normal distribution with mean 10 and standard deviation 3 use 

```{r}
rnorm(20, 10, 3)
```


```{r}
B <- 10000
x <- rnorm(B, 10, 3)
hplot(x, n=100)
```

The sample command above generally works for distributions that take finite discrete values with different probabilities. You can also generate many values in one step.

**Example**: say we have data from a normal distribution with mean 100 and standard deviation 25. What is the probability that an observation is less than 90?

```{r}
B <- 10000
x <- rnorm(B, 100, 25)
sum(x<90)/B
```

**Example:** say we roll a fair die 10 times. What is the mean and the standard deviation of the number of sixes?  

The number of "successes" in a sequence of Bernoulli trials is one of those special distributions. It is called the *Binomial* distribution and we can generate data from it with **rbinom(B, n, p)** where n is the number of trial and p is the success probability. So

```{r}
B <- 10000
x <- rbinom(B, 10, 1/6)
mean(x)
sd(x)
```

**Example:** In the beginning of the discussion on probability I had a table of probabilities. Let's recreate it!

Say we flip a fair coin 1000 times. What is the probability of 525 or more heads?


```{r}
B <- 10000
x <- rbinom(B, 1000, 1/2)
sum(x>=525)/B
```

Now we had that table for 505, .., 550, so:

```{r}
sum(x>=505)/B  
sum(x>=515)/B  
sum(x>=525)/B  
sum(x>=535)/B  
```

**Example:** At the end of the discussion on probability I had told you the formulas for the mean and the standard deviation of a Bernoulli trial: $\mu=p$ and $\sigma=\sqrt{p(1-p)}$. Let's test that!

```{r}
p <- 0.5
B <- 10000
x <- sample(0:1, size=B, replace=TRUE, prob=c(1-p, p))
round(c(p, mean(x)), 4)
round(c(sqrt(p*(1-p)), sd(x)), 4)
```

```{r}
p <- 0.15
B <- 10000
x <- sample(0:1, size=B, replace=TRUE, prob=c(1-p, p))
round(c(p, mean(x)), 4)
round(c(sqrt(p*(1-p)), sd(x)), 4)
```

```{r}
p <- 0.75
B <- 10000
x <- sample(0:1, size=B, replace=TRUE, prob=c(1-p, p))
round(c(p, mean(x)), 4)
round(c(sqrt(p*(1-p)), sd(x)), 4)
```
